Untitled

### 2022-12-16 draft_report.Rmd
updated scripts

source(paste0(root,"/haplotype_reconstruction_pipelines.R"))

@@

```{r}
cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

```{r}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

```{r}
f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = F, samples_gen = F)
```

@@

### GigaMUGA mutant
gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'GigaMUGA'
html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
summary_df <- summary_df[summary_df$`% No Call` < 10,]

@@

```{r}
### MiniMUGA pipeline
minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'MiniMUGA'
html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
minimuga_results <- list()
mini_pct_res <- list()
for (ind in ind_mini) {
  #ind <- 'JXN'
  print(ind)
  sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F)
  df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

@@

library(ggplot2)    
ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('qtl2 - proportion concordance between optimal and implemented models')
```

According to the visualization, which shows the percentage of marker differences between qtl2 genail and the truth models - DO, F3 and F2 are all producing results that are almost exactly the same as the optimal model, with CC having some discrepancies. We believe the reason why CC performs so differently is that the optimal model for CC does not allow any heterozygosity, whereas the genail does, so the genail model showed some residual heterozygosity from the CC mice which the optimal model does not have the ability to show.

@@

library(ggplot2)    
ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa - proportion concordance between optimal and implemented models')
```

@@

library(ggplot2)    
ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('proportion concordance between implemented models (haploqa vs qtl2)')
```

@@

(add screenshots of karyotype plots for outliers)

@@

(make a plot to visualize run time)
(talk about the strengths and weaknesses)
(figure out the error LOD)
qtl2 does not scale well with large number of founders

@@

qtl2fst - store large objects 
calc_genoprob_fst
(allele prob fst)
for DO - 
time the regular genoprob and fst
get the file size/memory usage of RDS and FST

>>>>

source(paste0(root,"/haplotype_reconstruction_pipelines.R"))
results_dir <- file.path(root, 'results')

@@

```{r message=FALSE}
cc_results <- haplotype_reconstruction_pipeline('CC', list_pheno, qtl2_file_gen = F, samples_gen = F)

@@

```{r}
bxd_results <- haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)

@@

```{r}
f2_results <- haplotype_reconstruction_pipeline('F2', list_pheno, qtl2_file_gen = T, samples_gen = F)
```

@@

```{r}
### GigaMUGA mutant
gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'GigaMUGA'
html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
summary_df <- summary_df[summary_df$`% No Call` < 10,]

@@

```{r}
### MiniMUGA pipeline
minimuga_url <- 'http://haploqa-dev.jax.org/tag/MiniMUGA.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'MiniMUGA'
html_table <- read_html(minimuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(minimuga_url), url_list, marker_type)
ind_mini <- summary_df[summary_df$`Haplotype Candidate` == 'False',]$`ID`
ind_mini <- ind_mini[ind_mini!='JXX'] # screentime error, skip this one
ind_mini <- ind_mini[ind_mini!='JY9'] # only 1 contributing strain, AIL incompatible, skip this one as well
ind_mini <- ind_mini[ind_mini!='JY6'] # temporary - JXU, JY3, JY5
minimuga_results <- list()
mini_pct_res <- list()
for (ind in ind_mini) {
  #ind <- 'JXN'
  print(ind)
  sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
  df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

@@

library(ggplot2)    
ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('qtl2 - proportion discordance between optimal and implemented models')
```

According to the visualization, which shows the percentage of marker discordance between qtl2 genail and the truth models - DO, F3 and F2 are all producing results that are almost exactly the same as the optimal model, with CC having some discrepancies. We believe the reason why CC performs so differently is that the optimal model for CC does not allow any heterozygosity, whereas the genail does, so the genail model showed some residual heterozygosity from the CC mice which the optimal model does not have the ability to show.

The outliers at 80% discordance are samples V6 and VP. Their haplotype reconstruction results appear to be highly heterozygous as CC mice, which explains the high discordance rate. However, it is believed that these two mice samples may have been mistakenly labeled as CC, as the genomes of CC mice are generally homozygous. Further investigation may be needed on the two samples.

Below are the karyotype plots for the two samples:

![Karyotype plot - sample VP in CC](/Users/linc/Desktop/VP_geno_cc.png) ![Karyotype plot - sample V6 in CC](/Users/linc/Desktop/V6_geno_cc.png)

@@

library(ggplot2)    
ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa - proportion discordance between optimal and implemented models')
```

@@

library(ggplot2)    
ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('haploqa vs qtl2 - proportion discordance between implemented models')
```

@@

@@

Aside from the statistical metrics comparison between the haplotype reconstruction results, it is also worth taking the programmatic strengths and weaknesses of the two models into consideration:

HaploQA:
  1. advantages: 
      a. Not a programming tool - people without relevant knowledge can still use the tool and perform haplotype reconstruction
      b. Works for all MUGA array-based genotyping data 
      c. Detailed and concise visualizations on web interface
  2. disadvantages:
      a. haplotype reports are not easy to access computationally
      b. source codes are also not easy to access computationally
      c. expensive to maintain at the organization's costs

R/qtl2: 
  1. advantages:
      a. fully accessible computationally
      b. capable of performing other genetic analysis
      c. currently maintained by developers
  2. disadvantages:
      a. algorithms is probability based, tends to bounce states when low-contribution founders are present
      b. does not scale well with large number of founders, run time increases exponentially 
      c. visualizations are less detailed and need to be generated on demand

@@
